### 2019-04-02 1_introduction.tex
Small tweaks.

The challenge of parallel and distributed computation has led to a
wide variety of proposals for programming models, languages, and
runtime systems. But while these systems are well-represented in the literature, comprehensive and comparative performance evaluations
remain difficult to find. Our
goal in this paper is to develop a meaningful and useful framework for
comparing the performance of parallel and distributed programming
systems.

@@

One approach to comparing the performance of different systems is through \emph{proxy-} or
\emph{mini-apps}. Because they distill the structure of a larger
application, mini-apps can provide insight
without the expense of developing a production code. However, despite the name, our experience is that
mini-apps still requite significant investment to develop
to the level of quality needed for useful benchmarking. In many cases,
the effort to tune for performance exceeds the effort to develop the
mini-app in the first place. As a result, implementations of mini-apps
often do not reach the level of maturity required to compare
systems. Indeed, there are few published studies
that involve more than a handful of systems~\cite{LULESH13}.

There is a second important issue: How does one predict how
the performance of a mini-app generalizes to a full
application, even one closely related? Part of the problem is
that \emph{weak} and \emph{strong} scaling, which are by far the most commonly
reported performance metrics,
do not directly characterize the performance of the underlying
programming system. Weak scaling can hide arbitrary amounts of runtime
system overhead with sufficiently large problem sizes, and strong
scaling does not separate runtime system overhead from other costs
(such as increased communication, among many other things) when
using progressively larger portions of a machine.

@@

To address the second issue and isolate the contribution of the programming model's runtime system
to application performance, we introduce a new metric called
\emph{minimum effective task granularity} (METG). Intuitively,
METG(50\%) is the smallest application task granularity which can be
executed while maintaining at least 50\% efficiency,
meaning that the application achieves at least 50\% of peak
performance (e.g., FLOPS or memory bandwidth) 
on a given machine. METG captures the essence of what is important in a weak
or strong scaling study, the behavior at the limit of scalability. For
weak scaling, METG(50\%) corresponds directly to the smallest problem
size that can be weak scaled with 50\% efficiency. For strong scaling,
METG(50\%) can be used to compute the scale at which performance can
be expected to dip below 50\% efficiency.

A lower METG does not necessarily mean that performance for a particular workload will be
better. For example, two systems with METG(50\%) of 100~us and 1~ms,
respectively, running an application with an average task granularity
of 10~ms, are both very likely to perform well. Only when the task
granularity approaches (or drops below) METG(50\%) will performance
likely diverge. Thus METG helps to identify the regime in which a
given system can deliver good performance.

Task Bench and METG address issues common in limit studies of runtime
systems for parallel and distributed programming. Such studies often
employ the metric \emph{tasks per second} (TPS), which is almost
universally measured with trivial (i.e., no) dependencies \cite{LegionTracing18, Canary16, Armstrong14}. While
phrased in terms of tasks, TPS can be measured for any system as long
as the application in question has identifiable units of work that run
to completion without interruption. TPS is an upper bound on
runtime-limited application throughput, but not a tight bound, as the
cost of non-trivial dependencies can be significant and may impact
scalability. This issue is easily fixed by running non-trivial
configurations of Task Bench.

>>>>

The challenge of parallel and distributed computation has led to a
wide variety of proposals for programming models, languages, and
runtime systems. While these systems are well-represented in the literature, comprehensive and comparative performance evaluations
remain difficult to find. Our
goal in this paper is to develop a meaningful and useful framework for
comparing the performance of parallel and distributed programming
systems.

@@

One approach to comparing the performance of different systems is through \emph{proxy-} or
\emph{mini-apps}. Because they distill the structure of a larger
application, mini-apps can provide insight
without the expense of developing a production code. However, despite the name, our experience is that
mini-apps still require significant investment to develop
to the level of quality needed for useful benchmarking. In many cases,
the effort to tune for performance exceeds the effort to develop the
mini-app in the first place. As a result, implementations of mini-apps
often do not reach the level of maturity required to compare
systems. Few published studies involve more than a handful of systems~\cite{LULESH13}.

There is a second important issue: How does one predict how
the performance of a mini-app generalizes to a full
application, even one closely related? Part of the problem is
that \emph{weak} and \emph{strong} scaling, by far the most commonly
reported performance metrics,
do not directly characterize the performance of the underlying
programming system. Weak scaling can hide arbitrary amounts of runtime
system overhead with sufficiently large problem sizes, and strong
scaling does not separate runtime system overhead from other costs
(such as increased communication, among many other things) when
using progressively larger portions of a machine.

@@

To address the second issue and isolate the contribution of the programming model's runtime system to application performance, we introduce a new metric called \emph{minimum effective task granularity} (METG). Intuitively, METG(50\%) is the smallest executable task granularity that maintains at least 50\% efficiency, meaning that the application achieves at least 50\% of peak
performance (e.g., FLOPS or memory bandwidth) on a given machine. METG captures the essence of what is important in a weak
or strong scaling study, the behavior at the limit of scalability. For
weak scaling, METG(50\%) corresponds directly to the smallest problem size that can be weak-scaled with 50\% efficiency. For strong scaling, METG(50\%) can be used to compute the scale at which performance can be expected to dip below 50\% efficiency.

A lower METG does not necessarily mean that performance for a particular workload is
better. For example, two systems with METG(50\%) of 100~us and 1~ms,
respectively, running an application with an average task granularity
of 10~ms, are both very likely to perform well. Only when the task
granularity approaches (or drops below) METG(50\%) will performance
likely diverge. Thus METG helps to identify the regime in which a
given system can deliver good performance.

Task Bench and METG address issues common in limit studies of runtime
systems for parallel and distributed programming. Such studies often
employ the metric \emph{tasks per second} (TPS), which is almost
universally measured with trivial (i.e., no) dependencies \cite{LegionTracing18, Canary16, Armstrong14}. While
phrased in terms of tasks, TPS can be measured for any system as long
as the application in question has identifiable units of work that run
to completion without interruption. TPS is an upper bound on
runtime-limited application throughput, but not a tight bound, as the
cost of non-trivial dependencies can be significant and may impact
scalability. This issue can be easily fixed by running non-trivial
configurations of Task Bench.
